Relative parabolic factorizations of Chevalley groups

### 2015-04-28 stability-conditions.tex
Small fixes
Relative ASR Bak-like

\subsection{Absolute stable rank}
For a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ denote by $\mathfrak{J}(a)$ the intersection of all maximal ideals of $R$ containing $a_1,\ldots,a_n$.
It is easy to see that a row $a\in R^n$ is unimodular if and only if $\mathfrak{J}(a)=R$. 
Clearly, for any $g\in\GL(n,R)$ one has $\mathfrak{J}(a\cdot g)=\mathfrak{J}(a)$.
\begin{dfn}
We say that a pair $(R,I)$ satisfies $\ASR_n(R,I)$ if it satisfies $\SR_n(R,I)$ and, moreover, 
for any row $a=(a_1,\ldots,a_{n+1})\in I^n$ there exist $b_1,\ldots,b_n\in I$ such that
\[\mathfrak{J}(a_1,\ldots,a_{n+1})=\mathfrak{J}(a_1+b_1a_{n+1},\ldots,a_n+b_na_{n+1}).\]
It is easy to see that $\ASR_m(R,I)$ implies $\ASR_n(R,I)$ for any $n\geqslant m$. 
By definition, the \emph{absolute relative stable rank} $\asr(R,I)$ is the smallest natural $n$ such that $\ASR_n(R, I)$ holds. 
Similarly, we set $\asr(R,I)=\infty$ if no such $n$ exists.
When $I=R$ we call $\asr(R, R)$ simply the \emph{absolute stable rank} of $R$ and denote it by $\asr(R)$.
\end{dfn}
The following reformulation of the absolute stable rank condition was noted by A.~Bak.
\begin{lemma} For a commutative ring $R$ the following statements are equivalent:
\begin{enumerate}
 \item\label{asr_Stein} $R$ satisfies $\ASR_n(R, R)$;
 \item\label{asr_Bak} For any \emph{unimodular} row $(a_1,\ldots,a_n,b,d)\in\Ums(n+2,R)$ there exist $c_1,\ldots,c_n\in R$ such that $(a_1+c_1d,\ldots,a_n+c_nd,b+td)$ is unimodular for all $t\in R$.
\end{enumerate}
\end{lemma}
\begin{proof}
Assume first that $\ASR_n(R)$ holds.
Then for the row $(a_1,\ldots,a_n,d)$ there exist $c_1,\ldots,c_n$ such that $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a_1+c_1d,\ldots,a_n+c_nd)$. 
Therefore $a_1+c_1d,\ldots,a_n+c_nd$ together with $b$ generate the unit ideal.
Of course, one could replace $b$ with $b+td$ from the very start.

To show the converse take arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!R$ and consider the unimodular row $(a_1,\ldots,a_n,1,d)\in\Ums(n+2,R).$
There exist $c_1,\ldots,c_n$ such that \[(a'_1,\ldots,a'_n,1+td)=(a_1+c_1d,\ldots,a_n+c_nd,1+td)\] is unimodular for all $t\in R$.
Suppose that there exists a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but at least one of the elements $d$, $a_i$ is not in $\mathfrak{m}$.
Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$ contrary to the assumption).
On the other hand, $1+td\notin\mathfrak{m}$,
hence the inequality $1+\bar{t}\bar{d}\neq0$ in the residue field $R/\mathfrak{m}$ may not hold for arbitrary $\bar{t}$ 
(take e.g. $\bar{t}=-1/\bar{d}$).
This shows that no such $\mathfrak{m}$ may exist and, therefore, $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a'_1,\ldots,a'_n)$.
\end{proof}

@@

The next lemma is an immediate consequence of the definition of stable rank.
\begin{lemma}\label{lemma:srUnip} Let $v\in\Ums(n, R, I)$ be a unimodular column. Assume that $\sr(I)\leq n-m$ for some $1\leq m \leq n$.
Then one can find a matrix $a\in \M(n-m,m,I)$ such that the column $(e_{n-m},a)v\in \Ums(n-m, R, I)$. \end{lemma}

>>>>

\subsection{Absolute stable rank}
For a row $a=(a_1,\ldots,a_n)\in{}^n\!R$ denote by $\mathfrak{J}(a)$ the intersection of all maximal ideals of $R$ containing $a_1,\ldots,a_n$.
It is easy to see that a row $a\in R^n$ is unimodular if and only if $\mathfrak{J}(a)=R$. 
Clearly, for any $g\in\GL(n,R)$ one has $\mathfrak{J}(a\cdot g)=\mathfrak{J}(a)$.
\begin{dfn}\label{dfn:j-stable}
We say that a row $a=(a_1,\ldots,a_{n+1})\in{}^{n+1}\!I$ is $\mathfrak{J}$-stable, if there exist $c_1,\ldots,c_n\in I$ such that
\[\mathfrak{J}(a_1,\ldots,a_{n+1})=\mathfrak{J}(a_1+c_1a_{n+1},\ldots,a_n+c_na_{n+1}).\]
\end{dfn}
\begin{dfn}\label{dfn:asr}
We say that a pair $(R,I)$ satisfies $\ASR_n(R,I)$ if satisfies $\SR_n(R,I)$ and, moreover, any row $a\in{}^{n+1}\!I$ is $\mathfrak{J}$-stable.

It is easy to see that $\ASR_m(R,I)$ implies $\ASR_n(R,I)$ for any $n\geqslant m$. 
By definition, the \emph{absolute relative stable rank} $\asr(R,I)$ is the smallest natural $n$ such that $\ASR_n(R, I)$ holds. 
Similarly, we set $\asr(R,I)=\infty$ if no such $n$ exists.
When $I=R$ we call $\asr(R, R)$ simply the \emph{absolute stable rank} of $R$ and denote it by $\asr(R)$.
\end{dfn}
The following reformulation of the absolute stable rank condition was noted by A.~Bak.
\begin{lemma} For a commutative ring $R$ the following statements are equivalent:
\begin{enumerate}
 \item\label{asr_Stein} $R$ satisfies $\ASR_n(R, R)$;
 \item\label{asr_Bak} For any \emph{unimodular} row $(a_1,\ldots,a_n,b,d)\in\Ums(n+2,R)$ there exist $c_1,\ldots,c_n\in R$ such that $(a_1+c_1d,\ldots,a_n+c_nd,b+td)$ is unimodular for all $t\in R$.
\end{enumerate}
\end{lemma}
%\begin{proof}
%Assume first that $\ASR_n(R)$ holds.
%Then for the row $(a_1,\ldots,a_n,d)$ there exist $c_1,\ldots,c_n$ such that $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a_1+c_1d,\ldots,a_n+c_nd)$. 
%Therefore $a_1+c_1d,\ldots,a_n+c_nd$ together with $b$ generate the unit ideal.
%Of course, one could replace $b$ with $b+td$ from the very start.
%
%To show the converse take arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!R$ and consider the unimodular row $(a_1,\ldots,a_n,1,d)\in\Ums(n+2,R).$
%There exist $c_1,\ldots,c_n$ such that \[(a'_1,\ldots,a'_n,1+td)=(a_1+c_1d,\ldots,a_n+c_nd,1+td)\] is unimodular for all $t\in R$.
%Suppose that there exists a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but at least one of the elements $d$, $a_i$ is not in $\mathfrak{m}$.
%Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$ contrary to the assumption).
%On the other hand, $1+td\notin\mathfrak{m}$,
%hence the inequality $1+\bar{t}\bar{d}\neq0$ in the residue field $R/\mathfrak{m}$ may not hold for arbitrary $\bar{t}$ 
%(take e.g. $\bar{t}=-1/\bar{d}$).
%This shows that no such $\mathfrak{m}$ may exist and, therefore, $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a'_1,\ldots,a'_n)$.
%\end{proof}

We will use use the following version of the relative absolute stable rank condition.
\begin{lemma}\label{lemma:relative-asr-unimod}
For a commutative ring $R$ and its ideal $I$ the following two statements are equivalent:
\begin{enumerate}
\item\label{asr-j-stable} Any row $a\in{}^{n+1}\!I$ is $\mathfrak{J}$-stable;
\item\label{asr-bak-like} For any $I$-unimodular row $(b,a_1,\ldots,a_n,d)\in\Ums(n+s,R,I)$ there exist $c_1,\ldots,c_n\in I$ such that $(b+b',a_1+c_1d,\ldots,a_n+c_nd)$ is $I$-unimodular for any $b'\in J$, where $J=\langle a_1,\ldots,a_n,d \rangle$.
\end{enumerate}
\end{lemma}
\begin{proof}
Assume first that any row $a\in{}^{n+1}\!I$ is $\mathfrak{J}$-stable. In particular for a given unimodular row $(b,a_1,\ldots,a_n,d)\in\Ums(n+s,R,I)$ there exist $c_1,\ldots,c_n$ such that
\[\mathfrak{J}(a_1,\ldots,a_{n+1})=\mathfrak{J}(a_1+c_1a_{n+1},\ldots,a_n+c_na_{n+1}).\]
Therefore $(b,a_1+c_1d,\ldots,a_n+c_nd)$ is also unimodular. Of course, for any $b'\in J$ one could replace $b$ with $b+b'$ from the very start.

To show the converse take arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!I$ and consider the $I$-unimodular row $(1,a_1,\ldots,a_n,d)\in\Ums(n+2,R,I).$
There exist $c_1,\ldots,c_n$ such that
\[ (1+b',a'_1,\ldots,a'_n)=(1+b',a_1+c_1d,\ldots,a_n+c_nd) \]
is unimodular for any $b'\in J$.
Suppose that there exists a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but at least one of the elements $d$, $a_i$ is not in $\mathfrak{m}$.
Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$ contrary to the assumption).
Note that $1+b'\notin\mathfrak{m}$, so the equation $1+\bar{b}'=0$ can not hold in the residue field $R/\mathfrak{m}$. But for $t\in R$ such that $\bar{t}=-1/\bar{d}$ one can take $b'=td\in J$, a contradiction to $1+\bar{b}'\neq0$. Hence no such $\mathfrak{m}$ exists and $\mathfrak{J}(a'_1,\ldots,a'_n)=\mathfrak{J}(a_1,\ldots,a_n,d)$.
\end{proof}

@@

The next lemma is an immediate consequence of the definition of stable rank.
\begin{lemma}\label{lemma:srUnip} Let $v\in\Ums(n, R, I)$ be a unimodular column. Assume that $\sr(I)\leq n-m$ for some $1\leqslant m < n$.
Then one can find a matrix $a\in \M(n-m,m,I)$ such that the column $(e_{n-m},a)v\in \Ums(n-m, R, I)$. \end{lemma}
